### 2019-04-18 CV/cv-sections/honors.tex

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Honors \& Awards}

%------------------------------------------------

\begin{cvhonors}

\cvhonor
{Abramov’s Fund Scholarship}
{for top students of Moscow Institute of Physics and Technology}
{Moscow, Russia}
{2015-2017}

\cvhonor
{Honourable Mention} % Award
{APhO (Asian Physics Olympiad)} % Event
{Singapore} % Location
{2014} % Date(s)

\cvhonor
{Prize-Winner} % Award
{All-Russian School Physics Olympiad, Final Round} % Event
{Saint Petersburg, Russia} % Location
{2014} % Date(s)

\cvhonor
{Winner} % Award
{All-Russian School Programming Olympiad, Region Round} % Event
{Moscow, Russia} % Location
{2014} % Date(s)

\cvhonor
{Winner} % Award
{All-Russian School Math Olympiad, Region Round} % Event
{Moscow, Russia} % Location
{2014} % Date(s)

\cvhonor
{Russian President's Scholarship}
{for high school students}
{Russia}
{2012-2014}

\cvhonor
{Moscow Governor’s Scholarship}
{for high school students}
{Moscow, Russia}
{2012-2014}

\cvhonor
{Winner} % Award
{All-Russian School Physics Olympiad, Final Round} % Event
{Vladivostok, Russia} % Location
{2013} % Date(s)

\cvhonor
{Prize-Winner} % Award
{All-Russian School Physics Olympiad, Final Round} % Event
{Saransk, Russia} % Location
{2012} % Date(s)

\end{cvhonors}