### 2019-04-18 CV/cv-sections/talks.tex

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Conferences \& Talks}

%----------------------------------------------------------------------------------------
%	SECTION CONTENT
%----------------------------------------------------------------------------------------

\begin{cventries}

%------------------------------------------------

\cventry
{Talk: <<Stochastic Spectral and Conjugate Descent Methods>>} % Role
{Weekly seminar <<Automatic control and Optimization Theory>>} % Event
{Institute for Control Problems, Moscow, Russia} % Location
{26 Mar. 2019} % Date(s)
{ % Description(s)
	\begin{cvitems}
		\item {\href{http://www.mathnet.ru/php/seminars.phtml?option_lang=rus&presentid=23658}{Link 1}}
		\item {\href{https://www.ipu.ru/node/50794}{Link 2}}
	\end{cvitems}
}

\cventry
{Talk: <<Stochastic Distributed Learning with Gradient Quantization and Variance Reduction>>}
{Seminar <<Modern Optimization Methods>>}
{Moscow Institute of Physics and Technology, Moscow, Russia}
{25 Mar. 2019}
{
	\begin{cvitems}
		\item {\href{https://vk.com/miptopt?w=wall-132864820_40\%2Fall}{Link}}
	\end{cvitems}
}

\cventry
{Poster Presenter: <<Stochastic Spectral Descent Methods>>} % Role
{Traditional School (Control, Information and Otimization)} % Event
{Voronovo, Moscow Region, Russia} % Location
{Jun. 2018} % Date(s)
{ % Description(s)
	\begin{cvitems}
		\item {\href{https://sites.google.com/site/traditionalschool/}{School website 1}}
		\item {\href{http://ssopt.org}{School website 2}}
	\end{cvitems}
}

\cventry
{Poster Presenter: Stochastic Spectral Descent Methods} % Role
{Optimization and Big Data Workshop} % Event
{KAUST, Thuwal, Saudi Arabia} % Location
{Feb. 2018} % Date(s)
{ % Description(s)
	\begin{cvitems}
		\item {\href{https://obd.kaust.edu.sa/}{Workshop website}}
		\item {\href{https://epostersonline.com/obd2018/node/51}{Poster}}
	\end{cvitems}
}

%------------------------------------------------

\end{cventries}