### 2019-12-10 CV/cv-sections/talks.tex

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Conferences \& Talks}

%----------------------------------------------------------------------------------------
%	SECTION CONTENT
%----------------------------------------------------------------------------------------

\begin{cventries}

\cventry
{Talk: <<Revisiting Stochastic Extragradient Method>>} % Role
{\href{https://iccopt2019.berlin/index.php}{ICCOPT 2019}} % Event
{TU, Berlin, Germany} % Location
{5 Aug. 2019} % Date(s)
{
}

\cventry
{\href{https://2019.ds3-datascience-polytechnique.fr/wp-content/uploads/2019/06/DS3-608_2019.pdf}{Poster: <<Stochastic Distributed Learning with Gradient Quantization and Variance Reduction>>}} % Role
{\href{https://2019.ds3-datascience-polytechnique.fr}{Data Science Summer School 2019}} % Event
{École polytechnique, Paris, France} % Location
{26 Jun. 2019} % Date(s)
{
}

\cventry
{Poster: <<Stochastic Distributed Learning with Gradient Quantization and Variance Reduction>>} % Role
{\href{http://ssopt.org}{Traditional School (Control, Information and Otimization)}} % Event
{Voronovo, Moscow Region, Russia} % Location
{20 Jun. 2019} % Date(s)
{
}

\cventry
{Talk: <<Stochastic Spectral and Conjugate Descent Methods>>} % Role
{\href{https://www.ipu.ru/node/50794}{Weekly seminar <<Automatic control and Optimization Theory>>}} % Event
{IPU, Moscow, Russia} % Location
{26 Mar. 2019} % Date(s)
{
}

\cventry
{Talk: <<Stochastic Distributed Learning with Gradient Quantization and Variance Reduction>>}
{\href{https://vk.com/miptopt?w=wall-132864820_40\%2Fall}{Seminar <<Modern Optimization Methods>>}}
{MIPT, Moscow, Russia}
{25 Mar. 2019}
{
}

\cventry
{Poster: <<Stochastic Spectral Descent Methods>>} % Role
{\href{https://sites.google.com/site/traditionalschool/}{Traditional School (Control, Information and Otimization)}} % Event
{Voronovo, Moscow Region, Russia} % Location
{Jun. 2018} % Date(s)
{
}

\cventry
{\href{https://epostersonline.com/obd2018/node/51}{Poster: Stochastic Spectral Descent Methods}} % Role
{\href{https://obd.kaust.edu.sa/}{Optimization and Big Data Workshop}} % Event
{KAUST, Thuwal, Saudi Arabia} % Location
{Feb. 2018} % Date(s)
{
}

%------------------------------------------------

\end{cventries}