### 2020-11-18 CV/cv-sections/education.tex
CV small update

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Education}

%----------------------------------------------------------------------------------------
%	SECTION CONTENT
%----------------------------------------------------------------------------------------

\begin{cventries}

%------------------------------------------------

\cventry
{PhD in Computer Science } % Degree
{King Abdullah University of Science and Technology} % Institution
{Thuwal, Saudi Arabia} % Location
{Jan. 2020 - PRESENT} % Date(s)
{
}

\cventry
{MS in Computer Science } % Degree
{King Abdullah University of Science and Technology} % Institution
{Thuwal, Saudi Arabia} % Location
{Sep. 2018 - Dec. 2019} % Date(s)
{
}

\cventry
{BS in Applied Mathematics and Physics} % Degree
{Moscow Institute of Physics and Technology} % Institution
{Moscow, Russia} % Location
{Sep. 2014 - Jun. 2018} % Date(s)
{ % Description(s) bullet points
}

%------------------------------------------------

\end{cventries}