### 2021-02-27 CV/cv-sections/papers.tex
-

\cvsection{Papers}

\begin{cventries}

%\cventry
%{} % Authors
%{} % Title
%{} 
%{}
%{\href{}{}}

\cventry
{Dmitry Kovalev, Eduard Gorbunov, Elnur Gasanov, Peter Richtárik} % Authors
{Stochastic Spectral and Conjugate Descent Methods} % Title
{} 
{Feb 2020}
{\href{}{}}

\end{cventries}