### 2021-02-27 CV/cv-sections/papers.tex
CV

\cvsection{Papers}

\begin{cventries}

%\cventry
%{} % Authors
%{} % Title
%{\href{}{}} 
%{} % Date
%{}

\cventry
{Adil Salim, Laurent Condat, Dmitry Kovalev, Peter Richtárik} % Authors
{An Optimal Algorithm for Strongly Convex Minimization under Affine Constraints} % Title
{\href{https://arxiv.org/abs/2102.11079}{arXiv 2021}} 
{22 Feb 2021} % Date
{}

\cventry
{Dmitry Kovalev, Egor Shulgin, Peter Richtárik, Alexander Rogozin, Alexander Gasnikov} % Authors
{ADOM: Accelerated Decentralized Optimization Method for Time-Varying Networks} % Title
{\href{https://arxiv.org/abs/2102.09234}{arXiv 2021}} 
{18 Feb 2021} % Date
{}

\cventry
{Konstantin Mishchenko, Bokun Wang, Dmitry Kovalev, Peter Richtárik} % Authors
{IntSGD: Floatless Compression of Stochastic Gradients} % Title
{\href{https://arxiv.org/abs/2102.08374}{arXiv 2021}} 
{16 Feb 2021} % Date
{}

\cventry
{Alexander Rogozin, Alexander Beznosikov, Darina Dvinskikh, Dmitry Kovalev, Pavel Dvurechensky, Alexander Gasnikov} % Authors
{Decentralized Distributed Optimization for Saddle Point Problems} % Title
{\href{https://arxiv.org/abs/2102.07758}{arXiv 2021}} 
{15 Feb 2021} % Date
{}

\cventry
{Dmitry Kovalev, Anastasia Koloskova, Martin Jaggi, Peter Richtárik, Sebastian U. Stich} % Authors
{A Linearly Convergent Algorithm for Decentralized Optimization: Sending Less Bits for Free!} % Title
{AISTATS 2021} 
{03 Nov 2020} % Date
{}

\cventry
{Eduard Gorbunov, Dmitry Kovalev, Dmitry Makarenko , Peter Richtárik} % Authors
{Linearly Converging Error Compensated SGD} % Title
{\href{https://papers.nips.cc/paper/2020/hash/ef9280fbc5317f17d480e4d4f61b3751-Abstract.html}{NeurIPS 2020}} 
{23 Oct 2020} % Date
{}

\cventry
{Alexander Rogozin, Vladislav Lukoshkin , Alexander Gasnikov, Dmitry Kovalev, Egor Shulgin} % Authors
{Towards Accelerated Rates for Distributed Optimization over Time-varying Networks} % Title
{\href{https://arxiv.org/abs/2009.11069}{arXiv 2020}} 
{23 Sep 2020} % Date
{}

\cventry
{Dmitry Kovalev, Adil Salim, Peter Richtárik} % Authors
{Optimal and Practical Algorithms for Smooth and Strongly Convex Decentralized Optimization} % Title
{\href{https://papers.nips.cc/paper/2020/hash/d530d454337fb09964237fecb4bea6ce-Abstract.html}{NeurIPS 2020}} 
{21 Jun 2020} % Date
{}

\cventry
{Grigory Malinovsky, Dmitry Kovalev, Elnur Gasanov, Laurent Condat, Peter Richtárik} % Authors
{From Local SGD to Local Fixed Point Methods for Federated Learning} % Title
{\href{https://proceedings.icml.cc/paper/2020/hash/53fdae58e861476b182b0cd6beade809-Abstract.html}{ICML 2020}} 
{03 Apr 2020} % Date
{}

\cventry
{Zhize Li, Dmitry Kovalev, Xun Qian, Peter Richtárik} % Authors
{Acceleration for Compressed Gradient Descent in Distributed and Federated Optimization} % Title
{\href{https://proceedings.icml.cc/paper/2020/hash/18ad9899f3f21a5a1583584d5f11c0c0-Abstract.html}{ICML 2020}} 
{26 Feb 2020} % Date
{}

\cventry
{Dmitry Kovalev, Robert M. Gower, Peter Richtárik, Alexander Rogozin} % Authors
{Fast Linear Convergence of Randomized BFGS} % Title
{\href{https://arxiv.org/abs/2002.11337}{arXiv 2020}} 
{26 Feb 2020} % Date
{}

\cventry
{Filip Hanzely, Dmitry Kovalev, Peter Richtárik} % Authors
{Variance Reduced Coordinate Descent with Acceleration: New Method With a Surprising Application to Finite-Sum Problems} % Title
{\href{https://proceedings.icml.cc/paper/2020/hash/b571ecea16a9824023ee1af16897a582-Abstract.html}{ICML 2020}} 
{11 Feb 2020} % Date
{}

\cventry
{Sélim Chraibi, Ahmed Khaled, Dmitry Kovalev, Peter Richtárik, Adil Salim, Martin Takáč} % Authors
{Distributed Fixed Point Methods with Compressed Iterates} % Title
{\href{https://arxiv.org/abs/1912.09925}{arXiv 2019}} 
{20 Dec 2019} % Date
{}

\cventry
{Mohammad Alkousa, Darina Dvinskikh, Fedor Stonyakin, Alexander Gasnikov, Dmitry Kovalev} % Authors
{Accelerated methods for composite non-bilinear saddle point problem} % Title
{\href{https://link.springer.com/article/10.1134/S0965542520110020}{Computational Mathematics and Mathematical Physics 2020}} 
{09 Dec 2019} % Date
{}

\cventry
{Dmitry Kovalev, Konstantin Mishchenko, Peter Richtárik} % Authors
{Stochastic Newton and Cubic Newton Methods with Simple Local Linear-Quadratic Rates} % Title
{\href{https://sites.google.com/site/optneurips19/}{NeurIPS 2019 Workshop}} 
{03 Dec 2019} % Date
{}

\cventry
{Adil Salim, Dmitry Kovalev, Peter Richtárik} % Authors
{Stochastic Proximal Langevin Algorithm: Potential Splitting and Nonasymptotic Rates} % Title
{\href{https://papers.nips.cc/paper/8891-stochastic-proximal-langevin-algorithm-potential-splitting-and-nonasymptotic-rates}{NeurIPS 2019}} 
{28 May 2019} % Date
{}

\cventry
{Konstantin Mishchenko, Dmitry Kovalev, Egor Shulgin, Peter Richtárik, Yura Malitsky} % Authors
{Revisiting Stochastic Extragradient} % Title
{\href{http://proceedings.mlr.press/v108/mishchenko20a}{AISTATS 2020}} 
{27 May 2019} % Date
{}

\cventry
{Robert M. Gower, Dmitry Kovalev, Felix Lieder, Peter Richtárik} % Authors
{RSN: Randomized Subspace Newton} % Title
{\href{https://papers.nips.cc/paper/8351-rsn-randomized-subspace-newton}{NeurIPS 2019}} 
{26 May 2019} % Date
{}

\cventry
{Samuel Horváth, Dmitry Kovalev, Konstantin Mishchenko, Peter Richtárik, Sebastian U. Stich} % Authors
{Stochastic Distributed Learning with Gradient Quantization and Variance Reduction} % Title
{\href{https://arxiv.org/abs/1904.05115}{arXiv 2019}} 
{04 Apr 2019} % Date
{}

\cventry
{Dmitry Kovalev, Samuel Horváth, Peter Richtárik} % Authors
{Don't Jump Through Hoops and Remove Those Loops: SVRG and Katyusha are Better Without the Outer Loop} % Title
{\href{http://proceedings.mlr.press/v117/kovalev20a}{ALT 2020}} 
{24 Jan 2019} % Date
{}

\cventry
{Alexander Gasnikov, Dmitry Kovalev} % Authors
{A hypothesis about the rate of global convergence for optimal methods (Newton's type) in smooth convex optimization} % Title
{\href{http://www.mathnet.ru/php/archive.phtml?wshow=paper\&jrnid=crm\&paperid=253\&option\_lang=eng}{
						Computer Research\linebreak and Modeling  2018}} 
{28 Feb 2018} % Date
{}

\cventry
{Dmitry Kovalev, Eduard Gorbunov, Elnur Gasanov, Peter Richtárik} % Authors
{Stochastic Spectral and Conjugate Descent Methods} % Title
{\href{https://papers.nips.cc/paper/7596-stochastic-spectral-and-conjugate-descent-methods}{NeurIPS 2018}} 
{11 Feb 2018} % Date
{}

%\cventry
%{Dmitry Kovalev, Anastasia Koloskova, Martin Jaggi, Peter Richtarik, Sebastian U. Stich} % Role
%{A Linearly Convergent Algorithm for Decentralized Optimization: Sending Less Bits for Free!} % Title
%{} % Location
%{Nov. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/2011.01697}{arXiv 2020}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Eduard Gorbunov, Dmitry Kovalev, Dmitry Makarenko, Peter Richtárik} % Role
%{Linearly Converging Error Compensated SGD} % Title
%{} % Location
%{Oct. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://papers.nips.cc/paper/2020/hash/ef9280fbc5317f17d480e4d4f61b3751-Abstract.html}{NeurIPS 2020}}		
%	\end{cvitems}
%}	
%
%\cventry
%{Alexander Rogozin, Vladislav Lukoshkin, Alexander Gasnikov, Dmitry Kovalev, Egor Shulgin} % Role
%{Towards Accelerated Rates for Distributed Optimization over Time-varying Networks} % Title
%{} % Location
%{Sep. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/2009.11069}{arXiv 2020}}		
%	\end{cvitems}
%}	
%
%\cventry
%{Dmitry Kovalev, Adil Salim, Peter Richtárik} % Role
%{Optimal and Practical Algorithms for Smooth and Strongly Convex Decentralized Optimization} % Title
%{} % Location
%{Jun. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://papers.nips.cc/paper/2020/hash/d530d454337fb09964237fecb4bea6ce-Abstract.html}{NeurIPS 2020}}		
%	\end{cvitems}
%}	
%	
%	
%\cventry
%{Grigory Malinovsky, Dmitry Kovalev, Elnur Gasanov, Laurent Condat, Peter Richtárik} % Role
%{From Local SGD to Local Fixed Point Methods for Federated Learning} % Title
%{} % Location
%{Apr. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://proceedings.icml.cc/paper/2020/hash/53fdae58e861476b182b0cd6beade809-Abstract.html}{ICML 2020}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Zhize Li, Dmitry Kovalev, Xun Qian, Peter Richtárik} % Role
%{Acceleration for Compressed Gradient Descent in Distributed and Federated Optimization} % Title
%{} % Location
%{Feb. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://proceedings.icml.cc/paper/2020/hash/18ad9899f3f21a5a1583584d5f11c0c0-Abstract.html}{ICML 2020}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Dmitry Kovalev, Robert M. Gower, Peter Richtárik, Alexander Rogozin} % Role
%{Fast Linear Convergence of Randomized BFGS} % Title
%{} % Location
%{Feb. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/2002.11337}{arXiv 2020}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Filip Hanzely, Dmitry Kovalev, Peter Richtárik} % Role
%{Variance Reduced Coordinate Descent with Acceleration: New Method With a Surprising Application to Finite-Sum Problems} % Title
%{} % Location
%{Feb. 2020} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://proceedings.icml.cc/paper/2020/hash/b571ecea16a9824023ee1af16897a582-Abstract.html}{ICML 2020}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Sélim Chraibi, Ahmed Khaled, Dmitry Kovalev, Peter Richtárik, Adil Salim, Martin Takáč} % Role
%{Distributed Fixed Point Methods with Compressed Iterates} % Title
%{} % Location
%{Dec. 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/1912.09925}{arXiv 2019}}		
%	\end{cvitems}
%}	
%	
%\cventry
%{Dmitry Kovalev, Konstantin Mishchenko, Peter Richtárik} % Role
%{Stochastic Newton and Cubic Newton Methods with Simple Local Linear-Quadratic Rates} % Title
%{} % Location
%{Dec. 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://sites.google.com/site/optneurips19/}{NeurIPS 2019 Workshop: Beyond First Order Methods in Machine Learning}}
%	\end{cvitems}
%}	
%
%\cventry
%{Mohammad Alkousa, Darina Dvinskikh, Fedor Stonyakin, Alexander Gasnikov, Dmitry Kovalev} % Role
%{Accelerated methods for composite non-bilinear saddle point problem} % Title
%{} % Location
%{Dec. 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/1906.03620}{arXiv 2019}}
%	\end{cvitems}
%}	
%
%\cventry
%{Adil Salim, Dmitry Kovalev, Peter Richtárik} % Role
%{Stochastic Proximal Langevin Algorithm: Potential Splitting and Nonasymptotic Rates} % Title
%{} % Location
%{May 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://papers.nips.cc/paper/8891-stochastic-proximal-langevin-algorithm-potential-splitting-and-nonasymptotic-rates}{NeurIPS 2019}}
%	\end{cvitems}
%}	
%	
%\cventry
%{Konstantin Mishchenko, Dmitry Kovalev, Egor Shulgin, Peter Richtárik, Yura Malitsky} % Role
%{Revisiting Stochastic Extragradient} % Title
%{} % Location
%{May 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{http://proceedings.mlr.press/v108/mishchenko20a}{AISTATS 2020}}
%	\end{cvitems}
%}
%
%\cventry
%{Robert M. Gower, Dmitry Kovalev, Felix Lieder, Peter Richtárik} % Role
%{RSN: Randomized Subspace Newton} % Title
%{} % Location
%{May 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://papers.nips.cc/paper/8351-rsn-randomized-subspace-newton}{NeurIPS 2019}}
%	\end{cvitems}
%}
%
%\cventry
%{Samuel Horváth, Dmitry Kovalev, Konstantin Mishchenko, Peter Richtárik, Sebastian U. Stich} % Role
%{Stochastic Distributed Learning with Gradient Quantization and Variance Reduction} % Title
%{} % Location
%{Jan. 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{https://arxiv.org/abs/1904.05115}{arXiv 2019}}
%	\end{cvitems}
%}
%
%\cventry
%{Dmitry Kovalev, Samuel Horváth, Peter Richtárik} % Role
%{Don't Jump Through Hoops and Remove Those Loops: SVRG and Katyusha are Better Without the Outer Loop} % Title
%{} % Location
%{Jan. 2019} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{http://proceedings.mlr.press/v117/kovalev20a}{ALT 2020}}
%	\end{cvitems}
%}
%
%
%\cventry
%{Alexander Gasnikov, Dmitry Kovalev	} % Role
%{A hypothesis about the rate of global convergence for optimal methods (Newton's type) in smooth convex optimization } % Title
%{} % Location
%{Feb. 2018} % Date(s)
%{ % Description(s)
%	\begin{cvitems}
%		\item {\href{http://www.mathnet.ru/php/archive.phtml?wshow=paper\&jrnid=crm\&paperid=253\&option\_lang=eng}{
%				Computer Research and Modeling, 2018,	Volume 10,	Issue 3,	Pages 305–314}}
%	\end{cvitems}
%}
%
%
%\cventry
%{Dmitry Kovalev, Eduard Gorbunov, Elnur Gasanov, Peter Richtárik} % Role
%{Stochastic Spectral and Conjugate Descent Methods} % Title
%{} % Location
%{Feb. 2018} % Date(s)
%{ % Description(s)
%\begin{cvitems}
%\item {\href{https://papers.nips.cc/paper/7596-stochastic-spectral-and-conjugate-descent-methods}{NeurIPS 2018}}
%\end{cvitems}
%}

\end{cventries}