### 2022-03-20 CV2/CV.tex
new cv

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plasmati Graduate CV
% LaTeX Template
% Version 1.0 (24/3/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro Plasmati (alessandro.plasmati@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Important note:
% This template needs to be compiled with XeLaTeX.
% The main document font is called Fontin and can be downloaded for free
% from here: http://www.exljbris.com/fontin.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

%!TEX TS-program = xelatex

\documentclass[a4paper,10pt]{article} % Default font size and paper size
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage{fontspec} % For loading fonts

%\setmainfont{[Fontin-Regular.otf]}
\defaultfontfeatures{Mapping=tex-text}
% \setmainfont{[Fontin-Regular.otf]} % Main document font
% \setmainfont[SmallCapsFont = Fontin-SmallCaps.otf]{[Fontin-Regular.otf]} % Main document font
\setmainfont{Fontin-Regular.otf}[
BoldFont       = Fontin-Bold.otf ,
ItalicFont     = Fontin-Italic.otf ,
SmallCapsFont  = Fontin-SmallCaps.otf ]

\usepackage{xunicode,xltxtra,url,parskip} % Formatting packages

\usepackage[usenames,dvipsnames]{xcolor} % Required for specifying custom colors

\usepackage[big]{layaureo} % Margin formatting of the A4 page, an alternative to layaureo can be \usepackage{fullpage}
% To reduce the height of the top margin uncomment: \addtolength{\voffset}{-1.3cm}

\usepackage{hyperref} % Required for adding links	and customizing them
\definecolor{linkcolour}{rgb}{0,0.2,0.6} % Link color
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour} % Set link colors throughout the document

\usepackage{titlesec} % Used to customize the \section command
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule] % Text formatting of sections
\titlespacing{\section}{0pt}{3pt}{3pt} % Spacing around sections

\usepackage{tabularx}

\begin{document}

\pagestyle{empty} % Removes page numbering

\font\fb=''[cmr10]'' % Change the font of the \LaTeX command under the skills section

%----------------------------------------------------------------------------------------
%	NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\par{\centering{\Huge Dmitry Kovalev}\bigskip\par} % Your name

\section{Personal Data}

\begin{tabular}{rl}
\textsc{Email} & \href{mailto:dakovalev1@gmail.com}{dakovalev1@gmail.com}\\
\textsc{Website} & \href{https://www.dmitry-kovalev.com}{www.dmitry-kovalev.com}\\
\textsc{Phone} & +7 905 719 06 98\\
\textsc{Google Scholar} &\href{https://scholar.google.com/citations?user=qHFA5z4AAAAJ}{https://scholar.google.com/citations?user=qHFA5z4AAAAJ}
\end{tabular}

%----------------------------------------------------------------------------------------
%	EDUCATION
%----------------------------------------------------------------------------------------

\section{Education}
\begin{tabular}{rl}	
\textsc{2014-2018}& BS in Applied Mathematics and Physics\\
& \textbf{Moscow Institute of Physics and Technology}, Dolgoprudny, Russia\\
& Advisor: \href{https://scholar.google.com/citations?user=AmeE8qkAAAAJ}{Alexander Gasnikov}\\
\textsc{2018-2019}& MS in Computer Science\\
& \textbf{King Abdullah University of Science and Technology}, Thuwal, Saudi Arabia\\
& Advisor: \href{https://richtarik.org}{Peter Richt\'{a}rik}\\
\textsc{2018-2021}& MS in Applied Mathematics and Physics\\
& \textbf{Moscow Institute of Physics and Technology}, Dolgoprudny, Russia\\
& Advisor: \href{https://scholar.google.com/citations?user=AmeE8qkAAAAJ}{Alexander Gasnikov}\\
\textsc{2019-Now}& PhD in Computer Science\\
& \textbf{King Abdullah University of Science and Technology}, Thuwal, Saudi Arabia\\
& Advisor: \href{https://richtarik.org}{Peter Richt\'{a}rik}\\
\end{tabular}

\section{Research Interests}
 Optimization Algorithms, Distributed Optimization, Machine Learning,

\section{Skills}

\begin{tabular}{rl}
	\textsc{Programming} & C/C++, Python; \textsc{Past experience:} Go, C\#, VB.NET\\
	\textsc{Computer} & macOS, LaTeX, Git\\
	\textsc{Mathematics} & Calculus, Linear Algebra, Probability and Statistics, Convex Analysis\\
\end{tabular}

%----------------------------------------------------------------------------------------
%	LANGUAGES
%----------------------------------------------------------------------------------------
\section{Languages}

\begin{tabular}{rl}
\textsc{English} & Advanced Knowledge \\
\textsc{Russian} & Mothertongue\\
\end{tabular}

\section{Awards}
\hypersetup{colorlinks,breaklinks,urlcolor=black,linkcolor=black}
\begin{tabular}{cl}
	\textsc{2021} & \textbf{CEMSE Research Excellence Award}, King Abdullah University of Science and Technology\\
	\textsc{2021} & \href{http://federated-learning.org/fl-icml-2021/}{\textbf{Best Student Paper Award}, FL-ICML 2021 Workshop}\\
	\textsc{2021} & \href{https://yandex.ru/scholarships/scholars}{\textbf{Ilya Segalovich Scientific Prize}, Yandex} \\
	\textsc{2020} & \href{https://yandex.ru/scholarships/scholars}{\textbf{Ilya Segalovich Scientific Prize}, Yandex}\\
	\textsc{2015-2017} & \textbf{Abramov’s Fund Scholarship}, Moscow Institute of Physics and Technology\\
	\textsc{2014} & \textbf{Asian Physics Olympiad (APhO), Honourable Mention}, Singapore\\
	\textsc{2014} & \textbf{All-Russian School Physics Olympiad, Final Round Prize-Winner}, Saint-Petersburg\\
	\textsc{2014} & \textbf{All-Russian School Programming Olympiad, Region Round Winner}, Moscow\\
	\textsc{2014} & \textbf{All-Russian School Math Olympiad, Region Round Winner}, Moscow\\
	\textsc{2012-2014} & \textbf{Russian President’s Scholarship for High School Sudents}\\
	\textsc{2012-2014} & \textbf{Moscow Governor’s Scholarship for High School Sudents}\\
	\textsc{2013} & \textbf{All-Russian School Physics Olympiad, Final Round Winner}, Vladivostok\\
	\textsc{2012} & \textbf{All-Russian School Physics Olympiad, Final Round Prize-Winner}, Saransk\\
\end{tabular}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}

\newpage

\section{Papers}

\input{papers.tex}

\end{document}