### 2022-03-23 CV2/general.tex

\section{Personal Data}

\begin{tabular}{rl}
	\textsc{Email} & \href{mailto:dakovalev1@gmail.com}{dakovalev1@gmail.com}\\
	\textsc{Website} & \href{https://www.dmitry-kovalev.com}{www.dmitry-kovalev.com}\\
	\textsc{Phone} & +7 905 719 06 98\\
	\textsc{Google Scholar} &\href{https://scholar.google.com/citations?user=qHFA5z4AAAAJ}{https://scholar.google.com/citations?user=qHFA5z4AAAAJ}
\end{tabular}

%----------------------------------------------------------------------------------------
%	EDUCATION
%----------------------------------------------------------------------------------------

\section{Education}
\begin{tabular}{rl}	
	\textsc{2014-2018}& BS in Applied Mathematics and Physics\\
	& \textbf{Moscow Institute of Physics and Technology}, Dolgoprudny, Russia\\
	& Advisor: \href{https://scholar.google.com/citations?user=AmeE8qkAAAAJ}{Alexander Gasnikov}\\
	\textsc{2018-2019}& MS in Computer Science\\
	& \textbf{King Abdullah University of Science and Technology}, Thuwal, Saudi Arabia\\
	& Advisor: \href{https://richtarik.org}{Peter Richt\'{a}rik}\\
	\textsc{2018-2021}& MS in Applied Mathematics and Physics\\
	& \textbf{Moscow Institute of Physics and Technology}, Dolgoprudny, Russia\\
	& Advisor: \href{https://scholar.google.com/citations?user=AmeE8qkAAAAJ}{Alexander Gasnikov}\\
	\textsc{2019-Now}& PhD in Computer Science\\
	& \textbf{King Abdullah University of Science and Technology}, Thuwal, Saudi Arabia\\
	& Advisor: \href{https://richtarik.org}{Peter Richt\'{a}rik}\\
\end{tabular}

\section{Research Interests}
%Optimization Algorithms, Distributed Optimization, Machine Learning
\begin{tabular}{cl}
	\qquad\qquad&  Optimization Algorithms\\
	&  Distributed Optimization\\
	&  Machine Learning
\end{tabular}

\section{Skills}

\begin{tabular}{rl}
	\textsc{Programming} & C/C++, Python; \textsc{Past experience:} Go, C\#, VB.NET, SQL, Julia\\
	\textsc{Computer} & macOS, LaTeX, Git\\
	\textsc{Mathematics} & Calculus, Linear Algebra, Probability and Statistics, Convex Analysis\\
\end{tabular}

%----------------------------------------------------------------------------------------
%	LANGUAGES
%----------------------------------------------------------------------------------------
\section{Languages}

\begin{tabular}{rl}
	\textsc{English} & Advanced Knowledge \\
	\textsc{Russian} & Mothertongue\\
\end{tabular}