### 2022-03-23 CV2/talks.tex

\section{Conferences and Talks}
\begin{enumerate}
\item {\bf Stochastic Spectral Descent Methods}, {\em Optimization and Big Data Workshop}, King Abdullah University of Science and Technology, Saudi Arabia (June 2018)
\item {\bf Stochastic Spectral Descent Methods}, {\em Traditional School (Control, Information and Optimization)}, Higher School of Economics Study Center, Voronovo, Rusia (February 2018)
\end{enumerate}