### 2022-03-23 CV2/talks.tex
talks

\section{Conferences and Talks}
\begin{enumerate}
\item {\bf Talk: Revisiting Stochastic Extragradient Method}, {\em International Conference on Continuous Optimization 2019}, Technical University, Berlin, Germany (August 2019)
\item {\bf Poster: Stochastic Distributed Learning with Gradient Quantization and Variance Reduction}, {\em Data Science Summer School 2019}, Ecole Polytechnique, Paris, France (June 2019)
\item {\bf Poster: Stochastic Distributed Learning with Gradient Quantization and Variance Reduction}, {\em Traditional School (Control, Information and Optimization)}, Higher School of Economics Study Center, Voronovo, Russia (June 2019)
\item {\bf Talk: Stochastic Spectral Descent Methods}, {\em Weekly Seminar <<Automatic control and Optimization Theory>>}, Institute for Control Problems, Moscow, Russia (March 2019)
\item {\bf Talk: Stochastic Distributed Learning with Gradient Quantization and Variance Reduction}, {\em Seminar <<Modern Optimization Methods>>}, Moscow Institute of Physics and Technology, Moscow, Russia (March 2019)
\item {\bf Poster: Stochastic Spectral Descent Methods}, {\em Optimization and Big Data Workshop}, King Abdullah University of Science and Technology, Thuwal, Saudi Arabia (June 2018)
\item {\bf Poster: Stochastic Spectral Descent Methods}, {\em Traditional School (Control, Information and Optimization)}, Higher School of Economics Study Center, Voronovo, Russia (February 2018)
\end{enumerate}