### 2022-03-24 CV2/teaching.tex
-

\section{Teaching Experience}

\newcommand{\teaching}[3]{{\bf #1}, #2 (#3)}

\begin{enumerate}
	\item \teaching{Mentor for 2 Research Projects with Students: Grigory Malinovsky and Abdurakhmon Sadiev}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{March 2022}

%	Mini-seminar for

	\item \teaching{Project Mentor at <<Modern Information, Optimization and Control Methods>> Student Educational Program}{Sirius University, Sochi, Russia}{July-August 2021}

	\item \teaching{Mentor for a Research Project with Student Grigory Malinovsky (resulted in ICML 2020 paper <<From Local SGD to Local Fixed Point Methods for Federated Learning>>)}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{January 2020}

%	 {\bf Project Mentor at <<Modern Information, Optimization and Control Methods>> Educational Program}, Sirius University, Sochi, Russia (July-August 2021)

\end{enumerate}