### 2022-03-27 CV2/teaching.tex
cv

\section{Teaching Experience}

\newcommand{\teaching}[3]{{\bf #1}, #2 (#3)}

\begin{enumerate}

	\item \teaching{Mentor for a Research Project with MIPT student Ekaterina Borodich (online)}{Moscow Institute of Physics and Technology, Moscow, Russia}{March 2022}

	\item \teaching{Mentor for 2 Research Projects with KAUST Students Grigory Malinovsky and Abdurakhmon Sadiev}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{March 2022}

	\item \teaching{Project Mentor at <<Modern Information, Optimization and Control Methods>> Student Educational Program}{Sirius University, Sochi, Russia}{July-August 2021}

	\item \teaching{Mentor for a Research Project with Student Grigory Malinovsky (done during his internship at KAUST, resulted in ICML 2020 paper <<From Local SGD to Local Fixed Point Methods for Federated Learning>>)}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{January 2020}

\end{enumerate}