### 2022-10-31 CV2/teaching.tex

\section{Teaching Experience}

\newcommand{\teaching}[3]{{\bf #1}, #2 (#3)}

\begin{enumerate}

	\item \teaching{Mentor for a Research Project with MIPT student Ekaterina Borodich (online, led to NeurIPS 2022 paper ``Optimal Gradient Sliding and its Application to Distributed Optimization Under Similarity'')}{Moscow Institute of Physics and Technology, Moscow, Russia}{March 2022}

	\item \teaching{Mentor for a Research Project with KAUST Student Abdurakhmon Sadiev (led to NeurIPS 2022 paper ``Communication Acceleration of Local Gradient Methods via an Accelerated Primal-Dual Algorithm with Inexact Prox'')}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{March 2022}

	\item \teaching{Project Mentor at <<Modern Information, Optimization and Control Methods>> Student Educational Program}{Sirius University, Sochi, Russia}{July-August 2021}

	\item \teaching{Mentor for a Research Project with Student Grigory Malinovsky (done during his internship at KAUST, led to ICML 2020 paper ``From Local SGD to Local Fixed Point Methods for Federated Learning'')}{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}{January 2020}

\end{enumerate}