### 2025-11-18 CV3/CV.tex
CV-ru

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plasmati Graduate CV
% LaTeX Template
% Version 1.0 (24/3/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro Plasmati (alessandro.plasmati@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Important note:
% This template needs to be compiled with XeLaTeX.
% The main document font is called Fontin and can be downloaded for free
% from here: http://www.exljbris.com/fontin.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

%!TEX TS-program = xelatex

\documentclass[a4paper,10pt]{article} % Default font size and paper size
\usepackage{fontspec} % For loading fonts

\defaultfontfeatures{Mapping=tex-text}

% \setmainfont{Fontin-Regular.otf}[
%   BoldFont       = Fontin-Bold.otf ,
%   ItalicFont     = Fontin-Italic.otf ,
% SmallCapsFont  = Fontin-SmallCaps.otf ]

\setmainfont{FontinSans_Cyrillic_R_46b.otf}[
  BoldFont       = FontinSans_Cyrillic_B_46b.otf,
  ItalicFont     = FontinSans_Cyrillic_I_46b.otf ,
  SmallCapsFont  = FontinSans_Cyrillic_SC_46b.otf
]
\usepackage{xunicode,xltxtra,url,parskip} % Formatting packages

\usepackage[usenames,dvipsnames]{xcolor} % Required for specifying custom colors

\usepackage[big]{layaureo} % Margin formatting of the A4 page, an alternative to layaureo can be \usepackage{fullpage}
% To reduce the height of the top margin uncomment: \addtolength{\voffset}{-1.3cm}

\usepackage{hyperref} % Required for adding links  and customizing them
\definecolor{linkcolour}{rgb}{0,0.2,0.6} % Link color
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour} % Set link colors throughout the document

\usepackage{titlesec} % Used to customize the \section command
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule] % Text formatting of sections
\titlespacing{\section}{0pt}{3pt}{3pt} % Spacing around sections

\usepackage{tabularx}

\begin{document}

\pagestyle{empty} % Removes page numbering

\font\fb=''[cmr10]'' % Change the font of the \LaTeX command under the skills section

%----------------------------------------------------------------------------------------
%  NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\par{\centering{\Huge Дмитрий Ковалев}\bigskip\par} % Your name

\input{general}

\input{awards.tex}

\input{papers.tex}

\input{talks.tex}

\input{teaching.tex}

\vspace{2em}
\input{date.tex}

\end{document}