### 2025-11-18 CV3/awards.tex

\section{Honors and Awards}
\hypersetup{colorlinks,breaklinks,urlcolor=black,linkcolor=black}
\begin{enumerate}
	\item  \textbf{CEMSE Student Research Excellence Award}, King Abdullah University of Science and Technology, 2021
	\item \href{http://federated-learning.org/fl-icml-2021/}{\textbf{Best Student Paper Award at FL-ICML 2021 Workshop}}
	\item \href{https://yandex.com/scholarships/#laureates}{\textbf{Ilya Segalovich Scientific Prize for Young Researchers 2021}, Yandex (highly selective: only 4 winners from Russia, Belarus, Kazakhstan)}
	\item {\bf  Annual PhD progress marked as Outstanding}, King Abdullah University of Science and Technology, 2019-2021
	\item \href{https://yandex.com/scholarships/#laureates}{\textbf{Ilya Segalovich Scientific Prize for Young Researchers 2020}, Yandex (highly selective: only 9 winners from Russia, Belarus, Kazakhstan)}
	\item {\bf Dean's Award}, given to top students accepted to King Abdullah University of Science and Technology, 2018
	\item \textbf{Abramov’s Fund Scholarship for Excellence in Study}, Moscow Institute of Physics and Technology, 2015-2017
	\item \textbf{Asian Physics Olympiad (APhO) 2014, Honourable Mention}, Singapore (participated as a member of Russian national team)

	\item \textbf{Russian President’s Scholarship for High School Sudents}, given for prize-winning at final round of All-Russian School Olympiads, 2012-2014
	\item \textbf{Moscow Governor’s Scholarship for High School Sudents}, given for prize-winning at region and final rounds of All-Russian School Olympiads, 2012-2014

	\item \textbf{All-Russian School Physics Olympiad, Final Round Prize-Winner}, Saint-Petersburg, 2014
	\item \textbf{All-Russian School Programming Olympiad, Region Round Winner}, Moscow, 2014
	\item \textbf{All-Russian School Math Olympiad, Region Round Winner}, Moscow, 2014

	\item \textbf{All-Russian School Physics Olympiad, Final Round Winner}, Vladivostok, 2013
	\item \textbf{All-Russian School Physics Olympiad, Final Round Prize-Winner}, Saransk, 2012
\end{enumerate}
%\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}